A supertree pipeline for summarizing phylogenetic and taxonomic information for millions of species

### 2017-01-25 doc/paper1-submission/paper5-adjusted.tex
Convert Algorithm to Figure.

When solving sub-problems, we sequentially incorporate splits from
trees in order of ranking, retaining splits that are compatible with
the current set of splits (Alg \ref{alg:ConsistentSplitsFromRankedList}).
The order of splits from the same tree is not specified by this approach,
and we incorporate splits using one of the possible post-order traversals
of the tree. We make use of the BUILD algorithm \citep{AhoSSU1981}
to assess compatibility. This strategy avoids unnecessary polytomies,
since splits of later input trees are only rejected from the summary
supertree if they conflict with higher-priority splits. Finally, we
use the BUILD algorithm to construct a supertree displaying all of
the splits in the set of compatible splits. Using the BUILD algorithm
to construct the subproblem summary tree satisfies criterion 3, because
trees from the BUILD algorithm do not contain unsupported branches.
\begin{algorithm}
\begin{algorithmic}
\Require An ordered list of $M$ splits, $\mathcal{R} = [R_1, R_2, R_3, \ldots, R3, \ldots, R_M]$
\State $\mathcal{C} = [R_1]$
\For{each split $i$ in $[2, 3 \ldots M]$}
   \State $\mathcal{T} \leftarrow \mathcal{C} + R_i$ \Comment{where `+' means concatenating 2 lists}
   \If{\textsc{BUILD}$(\mathcal{T})$ does not return null}
      \State $\mathcal{C} \leftarrow \mathcal{T}$
   \EndIf
\EndFor{}
\State\Return $\mathcal{C}$
\end{algorithmic}

\caption{\label{alg:ConsistentSplitsFromRankedList}ConsistentSplitsFromRankedList}
\end{algorithm}

@@

However, a taxon may be incompatible with the grafted solution; we
refer to such taxa as ``broken taxa.'' If a broken taxon contains
some unsampled children, it is not clear where these unsampled children
should be attached to the grafted solution. One approach would be
to mimic the application of Algorithm \ref{alg:ConsistentSplitsFromRankedList}
to the full (unpruned) taxonomic tree. This would be equivalent to
collapsing each edge in the taxonomy that attaches a broken taxon
to its parent. The unsampled children of broken taxa would attach
at their least inclusive ancestral taxon which is unbroken. In cases
of several adjacent taxa are broken, this can lead to polytomies of
very high degree deep in the tree. This can make the summary tree
difficult to navigate. Thus, we have adopted an alternative solution.
The \texttt{otc-unprune-solution-and-name-unnamed-nodes} tool from
otcetera attaches the unsampled children of a broken taxa to the grafted
solution as children of the MRCA of the sampled children.

>>>>

When solving sub-problems, we sequentially incorporate splits from
trees in order of ranking, retaining splits that are compatible with
the current set of splits (Fig. \ref{alg:ConsistentSplitsFromRankedList}).
The order of splits from the same tree is not specified by this approach,
and we incorporate splits using one of the possible post-order traversals
of the tree. We make use of the BUILD algorithm \citep{AhoSSU1981}
to assess compatibility. This strategy avoids unnecessary polytomies,
since splits of later input trees are only rejected from the summary
supertree if they conflict with higher-priority splits. Finally, we
use the BUILD algorithm to construct a supertree displaying all of
the splits in the set of compatible splits. Using the BUILD algorithm
to construct the subproblem summary tree satisfies criterion 3, because
trees from the BUILD algorithm do not contain unsupported branches.
\begin{figure}
\includegraphics[width=0.9\textwidth]{fig8}
\caption{\label{alg:ConsistentSplitsFromRankedList}Algorithm \texttt{ConsistentSplitsFromRankedList}}
\end{figure}

@@

However, a taxon may be incompatible with the grafted solution; we
refer to such taxa as ``broken taxa.'' If a broken taxon contains
some unsampled children, it is not clear where these unsampled children
should be attached to the grafted solution. One approach would be
to mimic the application of Algorithm $\texttt{ConsistentSplitsFromRankedList}$ (Fig. \ref{alg:ConsistentSplitsFromRankedList})
to the full (unpruned) taxonomic tree. This would be equivalent to
collapsing each edge in the taxonomy that attaches a broken taxon
to its parent. The unsampled children of broken taxa would attach
at their least inclusive ancestral taxon which is unbroken. In cases
of several adjacent taxa are broken, this can lead to polytomies of
very high degree deep in the tree. This can make the summary tree
difficult to navigate. Thus, we have adopted an alternative solution.
The \texttt{otc-unprune-solution-and-name-unnamed-nodes} tool from
otcetera attaches the unsampled children of a broken taxa to the grafted
solution as children of the MRCA of the sampled children.
